```{r}
#| echo: false
#| message: false
```

# Make (Yourself Look Good)

If you are part of a team that does code, then you already know the value of good documentation.

If you part of a team that does not code, then you may know the value of good documentation.

Apart from improving communication and avoiding frustration, good documentation makes you look good.

## Mind Maps (Architecture Overview)

A mind map is a diagram that shows the "from" and the "to", where we started and where we went. In data, it is often used for system architecture overviews: the eagle's eye look at the systems we use and how they interact. This is the data pipeline showcasing all its macro components (databases, Extract-Transfer-Load processes, APIs, cloud services, etc.).

DiagrammeR is the package we will use to draw our diagrams. Fortunately the language it employs, DOT, is fairly self-explanatory. (The name, however is, not. DOT stands for DAG of Tomorrow. DAG stands for Directed Acyclic Graph: a graph going in one direction and which does not loop back on itself.) There are no set rules for the colors and shapes (except to make them accessible).

```{r}
DiagrammeR::grViz("digraph {

graph [layout = dot, rankdir = LR]

# define the global styles of the nodes. We can override these in box if we wish
node [shape = rectangle, style = filled, fillcolor = Linen]

owner [label = 'Owner', shape = circle]
db [label = 'Network Data Store', shape = cylinder]
cloud [label = 'Cloud Data Service', shape = cylinder, fillcolor = lightblue]

datashape = cylinder

program [label = 'Application', fillcolor = pink]

code [label = 'Code Files', fillcolor = Beige]
file [label = 'Data Files']

}")
```

For big projects, these can get quite complicated... Just have a glance at the below.

```{r}
#| echo: false
example_complex <- DiagrammeR::grViz("digraph {

graph [layout = dot, rankdir = LR]

# define the global styles of the nodes. We can override these in box if we wish
node [shape = rectangle, style = filled, fillcolor = Linen]

fm [label = 'FM', shape = circle]
drive [label = 'Z Drive', shape = cylinder]
lake [label = 'Cloud Data Lake', shape = cylinder, fillcolor = lightblue]
cloud_etl [label = 'Cloud ETL Provider', shape = cylinder, fillcolor = lightblue]

network_api [label = 'Network API', shape = cylinder, fillcolor = lightblue]

datashape = cylinder

cad [label = 'AutoCAD', fillcolor = pink]
program [label = 'AutoLISP Script Writer', fillcolor = pink]

cs [label = 'CS', shape = circle]
op [label = 'OP', shape = circle]
tech [label = 'Tech.', shape = circle]

etl_shapes [label = 'Shapes ETL', fillcolor = Beige]
lisp [label =  'AutoLISP Script', fillcolor = Beige]
etl_wifi [label = 'Network ETL', fillcolor = Beige]

dxf [label =  '.dxf (Floor Plans)']
elev [label = 'Elevations']

dwg [label =  '.dwg (Floor Plans)']
drawings [label =  '.../Drawings', shape = folder]
data [label =  '/data', shape = folder]
logs [label = 'Network Logs']
data_occ [label = 'Network Data']
coord [label = 'Network Co-Ordinates']
data_shapes [label = 'Shapes Data']
data_app [label = 'App. Data']

{ rank = same; dxf; elev; coord; etl_shapes; etl_wifi; logs}
{ rank = same; dwg; cad; lisp}
{ rank = same; tech; drive}
{ rank = same; data_occ; data_shapes}
{ rank = same; network_api; cloud_etl; lake}
{ rank = same; cs; op}

# edge definitions with the node IDs
{fm} -> tech
{fm} -> drive

drive -> drawings -> dwg
drive -> data -> elev 
data -> program

cloud_etl -> etl_wifi
cloud_etl -> etl_shapes

lake -> logs
lake -> coord

fm -> etl_shapes
cs -> network_api
cs -> logs
op -> etl_wifi
etl_wifi -> data_occ
logs -> data_occ
network_api -> coord

cad -> dxf
tech -> program
tech -> dwg
tech -> cad
tech -> lisp
lisp -> dxf
program -> lisp
drawings -> lisp
tech -> elev
dwg -> dxf

cad -> dwg
dxf -> data_shapes
elev -> data_shapes
etl_shapes -> data_shapes
coord -> data_shapes

data_occ -> data_app
data_shapes -> data_app
}")
```

```{r}
#| eval: false
DiagrammeR::grViz("digraph {

graph [layout = dot, rankdir = LR]

# define the global styles of the nodes. We can override these in box if we wish
node [shape = rectangle, style = filled, fillcolor = Linen]

fm [label = 'FM', shape = circle]
drive [label = 'Z Drive', shape = cylinder]
lake [label = 'Cloud Data Lake', shape = cylinder, fillcolor = lightblue]
cloud_etl [label = 'Cloud ETL Provider', shape = cylinder, fillcolor = lightblue]

network_api [label = 'Network API', shape = cylinder, fillcolor = lightblue]

datashape = cylinder

cad [label = 'AutoCAD', fillcolor = pink]
program [label = 'AutoLISP Script Writer', fillcolor = pink]

cs [label = 'CS', shape = circle]
op [label = 'OP', shape = circle]
tech [label = 'Tech.', shape = circle]

etl_shapes [label = 'Shapes ETL', fillcolor = Beige]
lisp [label =  'AutoLISP Script', fillcolor = Beige]
etl_wifi [label = 'Network ETL', fillcolor = Beige]

dxf [label =  '.dxf (Floor Plans)']
elev [label = 'Elevations']

dwg [label =  '.dwg (Floor Plans)']
drawings [label =  '.../Drawings', shape = folder]
data [label =  '/data', shape = folder]
logs [label = 'Network Logs']
data_occ [label = 'Network Data']
coord [label = 'Network Co-Ordinates']
data_shapes [label = 'Shapes Data']
data_app [label = 'App. Data']

{ rank = same; dxf; elev; coord; etl_shapes; etl_wifi; logs}
{ rank = same; dwg; cad; lisp}
{ rank = same; tech; drive}
{ rank = same; data_occ; data_shapes}
{ rank = same; network_api; cloud_etl; lake}
{ rank = same; cs; op}

# edge definitions with the node IDs
{fm} -> tech
{fm} -> drive

drive -> drawings -> dwg
drive -> data -> elev 
data -> program

cloud_etl -> etl_wifi
cloud_etl -> etl_shapes

lake -> logs
lake -> coord

fm -> etl_shapes
cs -> network_api
cs -> logs
op -> etl_wifi
etl_wifi -> data_occ
logs -> data_occ
network_api -> coord

cad -> dxf
tech -> program
tech -> dwg
tech -> cad
tech -> lisp
lisp -> dxf
program -> lisp
drawings -> lisp
tech -> elev
dwg -> dxf

cad -> dwg
dxf -> data_shapes
elev -> data_shapes
etl_shapes -> data_shapes
coord -> data_shapes

data_occ -> data_app
data_shapes -> data_app
}")
```

```{r}
#| echo: false
#| column: margin
legend <- DiagrammeR::grViz("digraph {

graph [layout = dot, rankdir = LR]

# define the global styles of the nodes. We can override these in box if we wish
node [shape = rectangle, style = filled, fillcolor = Linen]

owner [label = 'Owner', shape = circle]
db [label = 'Network Data Store', shape = cylinder]
cloud [label = 'Cloud Data Service', shape = cylinder, fillcolor = lightblue]

datashape = cylinder

program [label = 'Application', fillcolor = pink]

code [label = 'Code Files', fillcolor = Beige]
file [label = 'Data Files']

}")
legend
```

```{r}
#| echo: false
example_complex
```

Phew, unless you need to join this project, you may not want to study this graph. But notice how the code is just the same amount of complicated? That's the beauty of the DOT language. Another beauty of this language is that it allows you to communicate a complicated system without sounding like the expert and single point-of-failure on the team. Finally, the complicated nature communicates the need for simplification and more support: E.g. why is there data that isn't in the lake? Why are there so many owners? These show the viewer that there are unoptimal systems in play and many players involved.

Below is a more sane example representing a completely automated system. Before you look, note:

- Qualtrics is a survey-creation and distribution software. 
- RStudio Server is the RStudio IDE on a server, rather than a desktop computer. 
- ES is an owner in the Environmental Safety department - they will design and send out the surveys. 
- CS is an owner in the Computing Services department - they will arrange permissions for the Microsoft Graph API which allows us to send emails using code.

```{r}
#| echo: false
example_simple <- DiagrammeR::grViz("digraph {

graph [layout = dot, rankdir = LR]

# define the global styles of the nodes. We can override these in box if we wish
node [shape = rectangle, style = filled, fillcolor = Linen]

drive [label = 'Z Drive', shape = cylinder]
server [label = 'Server', shape = cylinder]
db [label = 'Database', shape = cylinder]

wa [label = 'Work API', shape = cylinder, fillcolor = lightblue]
graph_api [label = 'Microsoft Graph API', shape = cylinder, fillcolor = lightblue]
qt_api [label = 'Qualtrics API', shape = cylinder, fillcolor = lightblue]

datashape = cylinder

qualtrics [label = 'Qualtrics', fillcolor = pink]
rstudio [label = 'RStudio Server', fillcolor = pink]

email [label = 'Email', fillcolor = lightblue]

es [label = 'ES', shape = circle]
da [label = 'Data Analyst', shape = circle]
it [label = 'Server Analyst', shape = circle]
cs [label = 'CS', shape = circle]

etl_workhours [label = 'Work-Hours ETL', fillcolor = Beige]
etl_inspections [label = 'Inspections ETL', fillcolor = Beige]

data [label =  '/data', shape = folder]
workhours [label =  '/work-hours', shape = folder]
workorder [label =  '/work-orders', shape = folder]
workhours_logs [label = 'Work-Hours ETL Logs']


cs -> graph_api
rstudio -> etl_workhours -> wa
da -> server -> rstudio
it -> server
etl_workhours -> workhours_logs

es -> qualtrics -> qt_api

qt_api -> graph_api -> email
wa -> graph_api
rstudio -> etl_inspections -> wa
etl_inspections -> qt_api
data -> workhours
data -> workorder
wa -> drive -> data
wa -> db
server -> db
}")
```

```{r}
#| eval: false
DiagrammeR::grViz("digraph {

graph [layout = dot, rankdir = LR]

# define the global styles of the nodes. We can override these in box if we wish
node [shape = rectangle, style = filled, fillcolor = Linen]

drive [label = 'Z Drive', shape = cylinder]
server [label = 'Server', shape = cylinder]
db [label = 'Database', shape = cylinder]

wa [label = 'Work API', shape = cylinder, fillcolor = lightblue]
graph_api [label = 'Microsoft Graph API', shape = cylinder, fillcolor = lightblue]
qt_api [label = 'Qualtrics API', shape = cylinder, fillcolor = lightblue]

datashape = cylinder

qualtrics [label = 'Qualtrics', fillcolor = pink]
rstudio [label = 'RStudio Server', fillcolor = pink]

email [label = 'Email', fillcolor = lightblue]

es [label = 'ES', shape = circle]
da [label = 'Data Analyst', shape = circle]
it [label = 'Server Analyst', shape = circle]
cs [label = 'CS', shape = circle]

etl_workhours [label = 'Work-Hours ETL', fillcolor = Beige]
etl_inspections [label = 'Inspections ETL', fillcolor = Beige]

data [label =  '/data', shape = folder]
workhours [label =  '/work-hours', shape = folder]
workorder [label =  '/work-orders', shape = folder]
workhours_logs [label = 'Work-Hours ETL Logs']


cs -> graph_api
rstudio -> etl_workhours -> wa
da -> server -> rstudio
it -> server
etl_workhours -> workhours_logs

es -> qualtrics -> qt_api

qt_api -> graph_api -> email
wa -> graph_api
rstudio -> etl_inspections -> wa
etl_inspections -> qt_api
data -> workhours
data -> workorder
wa -> drive -> data
wa -> db
server -> db
}")
```

```{r}
#| echo: false
#| column: margin
legend
```
```{r}
#| echo: false
example_simple
```

## Table Relationship Diagrams (AKA Entity Relationship Diagrams)

If you are working with multiple tables that relate to each other, you can improve communication with a diagram showing their primary and foreign keys.

Primary keys are columns that identify the unique rows. Foreign keys are columns that connect to primary keys in other tables.

For example, imagine you are working on a project that involves web-scraping countries' GDP per capita, life expectancy, corruption score, and more. Every measure is from a different web-page, meaning that you need to create tables for every measure: one table for GDP per capita, one for life expectancy, and so on. Finally, you want to compare this data based on when a country signed a human rights convention. This will be web-scraped from a United Nations web page. The ERD (Entity Relationship Diagram), or Data Model, will look like so:

```{r}
library(tibble)
library(dm)

## Creating fake, empty data
corruption <- tibble(
  country = NA,
  year = NA,
  corruption_score = NA
)

democracy <- tibble(
  country = NA,
  year = NA,
  democracy_score = NA
)

freedom <- tibble(
  country = NA,
  year = NA,
  freedom_index = NA
)

gdp <- tibble(
  country = NA,
  year = NA,
  gdp_per_capita = NA
)

safety <- tibble(
  country = NA,
  year = NA,
  safety_index = NA,
  crime_index = NA
)

unemployment <- tibble(
  country = NA,
  year = NA,
  unemployment_rate = NA
)

life_exp <- tibble(
  country = NA,
  year = NA,
  life_expectancy = NA
)

culture <- tibble(
  country = NA,
  year = NA,
  individualism = NA,
  indulgence = NA,
  power_distance = NA
)

un <- tibble(
  country = NA,
  date_signed_convention = NA
)

data_by_year <- tibble(
  country = NA,
  year = NA,
  corruption_score = NA,
  democracy_score = NA,
  freedom_index = NA,
  gdp_per_capita = NA,
  safety_index = NA,
  crime_index = NA,
  unemployment_rate = NA,
  life_expectancy = NA,
  individualism = NA,
  indulgence = NA,
  power_distance = NA
)

dm_one <- dm(
  corruption, 
  democracy,
  freedom,
  gdp,
  life_exp,
  safety,
  unemployment,
  culture,
  un
)

dm_by_source <- dm_one |> 
  ## Primary keys
  dm_add_pk(un, country) |> 
  dm_add_pk(corruption, c(country, year)) |> 
  dm_add_pk(democracy, c(country, year)) |> 
  dm_add_pk(gdp, c(country, year)) |> 
  dm_add_pk(life_exp, c(country, year)) |> 
  dm_add_pk(safety, c(country, year)) |> 
  dm_add_pk(unemployment, c(country, year)) |> 
  dm_add_pk(culture, c(country, year)) |> 
  dm_add_pk(freedom, c(country, year)) |> 
  ## Foreign keys
  dm_add_fk(corruption, country, un) |>
  dm_add_fk(democracy, country, un) |>
  dm_add_fk(gdp, country, un) |>
  dm_add_fk(life_exp, country, un) |>
  dm_add_fk(safety, country, un) |>
  dm_add_fk(unemployment, country, un) |>
  dm_add_fk(culture, country, un) |>
  dm_add_fk(freedom, country, un) |>
  ## Colors
  dm_set_colors(
    darkred = corruption,
    blue = freedom,
    gold = gdp,
    yellow = life_exp,
    darkgreen = safety, 
    brown = unemployment,
    
    whitesmoke = culture,
    pink = un
    ) |> 
  ## Draw from bottom to top
  dm_draw(view_type = "all", rankdir = "BT")

dm_by_source
```

Each year-dependent measure is in its own table, and its columns that identify unique rows are `country` and `year`. For the United Nations table, `country` is only needed to identify unique rows. All tables are thus connected by `country`.

We can join the bottom tables by country and year, to get us:

```{r}
dm(un, data_by_year) |> 
  ## Primary keys
  dm_add_pk(un, country) |> 
  dm_add_pk(data_by_year, c(country, year)) |> 
  ## Foreign keys
  dm_add_fk(data_by_year, country, un) |>
  dm_set_colors(
    darkred = data_by_year,
    pink = un
    ) |> 
  dm_draw(view_type = "all")
```

Furthermore, for added flair, we can re-arrange the data model based on whether higher values represent higher or lower quality of life:

```{r}
library(dplyr)
positive_qol <- Reduce(
  \(x,y) full_join(x, y, by = join_by(country, year)),
  list(freedom, democracy, gdp, life_exp, safety)
  ) |> 
  select(- crime_index)

negative_qol <- full_join(corruption, unemployment,
                          by = join_by(country, year))
negative_qol$crime_index <- NA

dm(positive_qol, negative_qol, un) |> 
  dm_add_pk(un, country) |> 
  dm_add_pk(negative_qol, c(country, year)) |> 
  dm_add_pk(positive_qol, c(country, year)) |> 
  dm_add_fk(negative_qol, country, un) |>
  dm_add_fk(positive_qol, country, un) |>
  dm_set_colors(
    darkred = negative_qol,
    blue = positive_qol,
    pink = un
  ) |> 
  dm_draw(view_type = "all", rankdir = "BT")

```



We can include the above code in a `.qmd` file and call it a day. But we can also create the diagram in a `.R` file (script) instead, export the diagram, and import the diagram in our `.qmd` file.

```{r}
#| eval: false
## In the .R file:
DiagrammeRsvg::export_svg(dm_by_source) |> write("output/documentation/dm_by_source.svg")

## In the .qmd file:
knitr::include_graphics("output/documentation/dm_by_source.svg")
```

